Comparing clustering parameters with rOpenScPCA

### 2025-01-13 analyses/hello-clusters/02_compare-clustering-parameters.Rmd
add section for varying multiple parameters, and other tweaks

## Perform clustering across a set of parameters

@@

## Visualizing clustering results

@@

```{r plot all umaps, fig.width = 12}
umap_plots <- cluster_results_list |>
  purrr::imap(
    \(cluster_df, clustering_name) {
      # Add a column with cluster assignments to umap_df
      umap_df_plot <- umap_df |>
        dplyr::mutate(cluster = cluster_df$cluster)

@@

## Evaluating clustering results

@@

### Silhouette width and neighborhood purity

@@

# Add together using the patchwork library, without a legend
silhouette_plot + purity_plot & theme(legend.position = "none")
```
While there does not appear to be a salient difference among silhouette width distributions, it does appear that purity is higher with a higher nearest neighbors parameter.

@@

>>>>

## Varying a single clustering parameter

@@

### Visualizing clustering results

@@

```{r plot nn umaps, fig.width = 12}
umap_plots <- cluster_results_list |>
  purrr::imap(
    \(cluster_df, clustering_name) {
      # Add a column with cluster assignments to umap_df
      umap_df_plot <- umap_df |>
        dplyr::mutate(cluster = cluster_df$cluster)

@@

### Evaluating clustering results

@@

#### Silhouette width and neighborhood purity

@@

# Add together using the patchwork library, without a legend
silhouette_plot + purity_plot & theme(legend.position = "none")
```

While there does not appear to be a salient difference among silhouette width distributions, it does appear that purity is higher with a higher nearest neighbors parameter.

#### Stability

@@

## Varying multiple clustering parameters

The previous section demonstrated how to calculate clusters and QC metrics when varying one parameter, but it is possible to vary multiple parameters at once with `rOpenScPCA::sweep_clusters()`.
In this section, we'll show an overview of how you might write code to vary two parameters at once (here, nearest neighbors and resolution as examples) and visualize results.

```{r sweep two parameters}
# Define vectors of parameters to vary
nn_values <- seq(10, 30, 10)
res_values <- seq(5, 15, 5) / 10

cluster_results_list <- rOpenScPCA::sweep_clusters(
  pca_matrix,
  nn = nn_values,
  resolution = res_values
)
```

This resulting list now has 9 different clustering results, for each combination of `nn_values` and `res_values`:

```{r length cluster_results_list two parameters}
length(cluster_results_list)
```

### Visualize clusters

Next, we'll iterate over `cluster_results_list` to plot the UMAPs.
This time, we'll use `purrr::map()` and pull out parameters from each iteration's `cluster_df` to form the UMAP panel title.

```{r plot nn res umaps, fig.height = 14}
umap_plots <- cluster_results_list |>
  purrr::map(
    \(cluster_df) {
      # Add a column with cluster assignments to umap_df
      umap_df_plot <- umap_df |>
        dplyr::mutate(cluster = cluster_df$cluster)

      # Create a title for the UMAP with both parameters
      umap_title <- glue::glue(
        "nn: {unique(cluster_df$nn)}; res: {unique(cluster_df$resolution)}"
      )

      # Plot the UMAP, colored by the new cluster variable
      ggplot(umap_df_plot, aes(x = UMAP1, y = UMAP2, color = cluster)) +
        geom_point(alpha = 0.6) +
        labs(title = umap_title) +
        # We'll add a couple UMAP-specific plot settings
        coord_equal() +
        theme(
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          legend.position = "bottom"
        )
    }
  )

### Calculate and visualize QC metrics

This section presents one coding strategy to calculate and visualize results when varying two clustering parameters.
In particular, we use faceting to help display all information in one plot, by placing nearest neighbor values on the X-axis and faceting by resolution values.
Since silhouette width and neighhorbood purity calculations using generally similar code, we'll just show neighborhood purity here.

#### Neighborhood purity

First, we'll calculate neighborhood purity and combine results into a single data frame.

```{r calculate purity two parameters}
purity_df <- cluster_results_list |>
  purrr::map(
    \(cluster_df) {
      rOpenScPCA::calculate_purity(pca_matrix, cluster_df)
    }
  ) |>
  dplyr::bind_rows()
```

We'll add a column `resolution_label` which we'll use to have informative panel titles in the faceted ggplot we make next.

```{r add resolution_label column}
purity_df <- purity_df |>
  dplyr::mutate(resolution_label = glue::glue("Resolution: {resolution}"))
```

```{r visualize purity two parameters}
ggplot(purity_df) +
  aes(x = as.factor(nn), y = purity, fill = as.factor(nn)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Pastel2") +
  # facet by resolution
  facet_wrap(vars(resolution_label)) +
  labs(
    x = "Number of nearest neighbors",
    y = "Neighborhood purity"
  ) +
  theme(legend.position = "none")
```

Similar to above, we'll calculate stability, combine results into a single data frame, add a `resolution_label` column to support plot interpretation, and finally make our plot.

```{r calculate stability two parameters}
stability_df <- cluster_results_list |>
  purrr::map(
    \(cluster_df) {
      # Extract parameters for this clustering result
      nn <- unique(cluster_df$nn)
      resolution <- unique(cluster_df$resolution)

      rOpenScPCA::calculate_stability(
        pca_matrix,
        cluster_df,
        nn = nn,
        resolution = resolution
      )
    }
  ) |>
  dplyr::bind_rows()

stability_df <- stability_df |>
  dplyr::mutate(resolution_label = glue::glue("Resolution: {resolution}"))
```

```{r visualize stability two parameters}
ggplot(stability_df) +
  aes(x = as.factor(nn), y = ari, fill = as.factor(nn)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Pastel2") +
  # facet by resolution
  facet_wrap(vars(resolution_label)) +
  labs(
    x = "Number of nearest neighbors",
    y = "Adjusted Rand Index"
  ) +
  theme(legend.position = "none")
```
